Constraining Quiescent Supermassive Black Hole Environments.

### 2014-09-02 ms.tex
Fixing typos in manuscript and commiting both appendices.

\newcommand{\pyear}{{\rm yr}^{-1}}
\newcommand{\tage}{t_{\rm age}}

@@

\author[Generozov, Metzger, \& Stone]{Aleksey Generozov$\thanks{E-mail: ag@astro.columbia.edu}$, Brian~D.~Metzger, Nicholas Stone$\footnotemark[1]$\\
Columbia Astropysics Labratory, Columbia University, 550 West 120th Street, New York, NY 10027}

@@

The location of $\rs$ and hence the accretion rate is quite sensitive
to the assumed heating rate.  Physically, the heating rate of the gas
would vary from galaxy to galaxy. For example, SNe Ia may be a large
source of energy for the gas. However, this will be only be the case
for larger black holes for which the gas inflow time is smaller than
the time between successive Ias. Thus, for smaller black holes, the
effective gas heating rate may be considerably smaller. Thus, we
investigate the effect of changing rate of energy injection on the gas
density profiles.

We compare our calculated gas profiles to those inferred from Chandra
X-ray observations by \citealt{AllenDunn+:2006a} and
\citealt{RussellMcNamara+:2013a}. Additionally, we chack how well our
results are able to reproduced relationships between $L_x$ and
$\Mbh$. In performing these comparisons we hope to determine how well
the gas environments of quiescent galaxies may be understood using the
relatively well understood physics of stellar winds and supernova
heating.
 %%AG--say something about x-ray observations.

@@

\section{Solutions}
\subsection{Analytics: Flow properties and $\rs$}
It is possible to derive an analytic expression for the steady-state stagnation radius, $\rs$, in terms of $v_w$ and the power law slope of the density (at $\rs$) $n$.   We summarize the main results here and place the detailed derivation in Appendix~\ref{app:rs}.

@@

We may also write an (implicit) analytic expression for $\rs$ itself. Let $\omega=\Mstar(\rs)/\Mbh$ and  $\eta=v_{w,0}/\sigma_{\rm soi}$.

\begin{align}
\x=\frac{1}{\eta^2 n}\left[ (1+\omega)\left(A-\frac{n}{2}\right)-B\right]
\label{eq:stag_analytic}
\end{align}

When $\Mstar << \Mbh$ at the stagnation radius, the relationship between $\rs$ and $\vw$ is greatly simplified.

@@

\[
\dot{M}=
\begin{cases}
   1.5 \times 10^{-3}  \left(\Mbheight\right)^{1.9} \left(\vwOFH\right)^{-4} \Msun \, \pyear& \text{for core galaxies} \\
   3.3 \times 10^{-3} \left(\Mbheight\right)^{1.4} \left(\vwOFH\right)^{-2}  \Msun \, \pyear  & \text{for cusp galaxies} 
\end{cases}
\label{eq:mdot_analytic}
\]

Finally, we may estimate the $\rho$ at $\rs$, using

\begin{equation}
\rho(\rs)\simeq q(\rs) \tff
\end{equation}

\subsection{Comparison with Bondi Accretion}
The Bondi accretion rate, $\Mdotb$ is given by

\begin{align}
\Mdotb=4\pi \lambda r_b^2 \rho(r_b) v_{\rm ff}(r_b)
\end{align}

where $r_b=G M/c_s^2$.

%%AG: I am not entirely comfortable with this approximation for the accretion rate. 
In our model the accretion rate may be approximated as the mass enclosed with $\rs$ divided by the free-fall time.

@@

Where have set $n=1$.  At $\rs$ we have

@@

\item \emph{Overall Heating Rate} The overall value of $\vwO$ has
  contributions from Type Ia SNes and main sequence winds.  The
  relative contributions will depend on both the age of the stellar
  population, $t_{\rm age}$ and $\Mbh$.  This is illustrated in
  Figure~\ref{fig:vweff}. For black hole masses below $10^7 \Msun$ SNe
  Ia will not contribute to $\vwO$ as the flow time scale is short
  compared to $t_{\rm Ia}$, and the heating is dominated by main
  sequence winds (with $\vwO\simeq 100$ km/s).  On the other hand, for
  $\Mbh \gsim 10^8 \Msun$, heating from Ias becomes important, as
  $\rs>\rIa$.  In this regime, we add the $\vwO$ values from SNe Ias
  and from main sequence winds in quadrature. $\vwO$ from SNe Ia's
  depends on the rate of mass loss ($\eta$), which is in turn a
  function of $t_{\rm age}$. This dependence is why the effective
  $\vwO$ increases with $\tage$ for large $\Mbh$.

@@

For small $\eta$ the effects of cooling will be less pronounced. For
each of our galaxies we give the maximum $\eta$ such that $H/C >$1 in
Table .  Roughly speaking, this is the maximum $\eta$ for which it
would be safe to ignore cooling. $\eta$ depends on the age of the
stellar population, and will realistically vary between $\sim 0.1$ and
$\sim 1$.

For $\vwO$=1000 km/s, it is apparent from Figure~\ref{fig:cooling}, that rad

@@

\appendix
\clearpage
\section{Analytic Expression for the stagnation radius}
\label{app:rs}
\input{appendix}

>>>>

\newcommand{\pyear}{{\rm yr}^{-1}}
\newcommand{\tage}{t_{\rm age}}
\renewcommand{\th}{t_h}

@@

\author[Generozov, Metzger, \& Stone]{Aleksey Generozov$\thanks{E-mail: ag@astro.columbia.edu}$, Brian~D.~Metzger, Nicholas Stone\\
Columbia Astropysics Labratory, Columbia University, 550 West 120th Street, New York, NY 10027}

@@

The location of $\rs$ and hence the accretion rate is quite sensitive
to the assumed heating rate.  Physically, the heating rate of the gas
would vary from galaxy to galaxy. For example, SNe Ia may be a large
source of energy for the gas. However, this will be only be the case
for larger black holes for which the gas inflow time is smaller than
the time between successive Ias. Thus, for smaller black holes, the
effective gas heating rate may be considerably smaller. We
investigate the effect of changing rate of energy injection on the gas
density profiles.

We compare our calculated gas profiles to those inferred from Chandra
X-ray observations by \citealt{AllenDunn+:2006a} and
\citealt{RussellMcNamara+:2013a}. Additionally, we chack how well our
results are able to reproduced relationships between $L_x$ and $\Mbh$
in \citealt{MillerGallo+:2014a}. In performing these comparisons we
hope to determine how well the gas environments of quiescent galaxies
may be understood using the relatively well understood physics of
stellar winds and supernova heating.
 %%AG--say something about x-ray observations.

@@

\section{Solutions}
\subsection{Analytics: Flow properties at $\rs$}
It is possible to derive an analytic expression for the steady-state stagnation radius, $\rs$, in terms of $v_w$ and the power law slope of the density (at $\rs$) $n$.   We summarize the main results here and place the detailed derivation in Appendix~\ref{app:rs}.

@@

We may also write an (implicit) analytic expression for $\rs$
itself. Let  $\eta=v_{w,0}/\sigma_{\rm soi}$. Then

\begin{align}
\x=\frac{1}{\eta^2 n}\left[ (1+\Mstar(\rs)/\Mbh)\left(A-\frac{n}{2}\right)-B\right]
\label{eq:stag_analytic}
\end{align}

Where A and B are are constants, which depend on the nuker $\gamma$
and the adiabatic index of the gas (see Appendix~\ref{app:rs})  $\Mstar << \Mbh$ at the stagnation radius, the relationship between $\rs$ and $\vw$ is greatly simplified.

@@

\begin{align}
\dot{M}\simeq
\begin{cases}
   1.5 \times 10^{-3}  \left(\Mbheight\right)^{1.9}
   \left(\vwOFH\right)^{-4} \eta \Msun \, \pyear& \text{for core galaxies} \\
   3.3 \times 10^{-3} \left(\Mbheight\right)^{1.4}
   \left(\vwOFH\right)^{-2} \eta \Msun \, \pyear  & \text{for cusp
     galaxies} 
 \end{cases}
\label{eq:mdot_analytic}
\end{align}

In our model the accretion onto the black hole may be approximated as
the gas mass enclosed inside $\rs$ divided by the free-fall time at $\rs$.

\begin{align}
&\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}(\rs)}
\label{eq:mdot_gas}
\end{align}

Finally, using Equations~\ref{eq:tstag} and \ref{eq:mdot_analytic}, we
may obtain a scaling relationship for $\rho(\rs)$.

\begin{align}
\rho(\rs)\simeq
\begin{cases}
2.5 \times 10^{-24} \Mbheight^{-0.14} \vwOFH^{-1} \eta \, {\rm g \, cm^{-3}}& \text{for   galaxies}\\
5.5 \times 10^{-24}  \Mbheight^{-0.56} \vwOFH    \eta \, {\rm g \,cm^{-3}} & \text{for cursp galaxies}
\end{cases}
\label{eq:rhors}
\end{align}

\subsection{Comparison with Bondi Accretion}
The Bondi accretion rate, $\Mdotb$ in terms of the Bondi radius $\rb=G
M/c_s^2$ is

\begin{align}
\Mdotb=4\pi \lambda r_b^2 \rho(r_b) v_{\rm ff}(r_b).
\end{align}

We may rewrite Equation~\ref{eq:mdot_gas} as follows

@@

Where have set the power law slope of the gas density, $n$, to 1.  At $\rs$ we have

@@

\item \emph{Overall Heating Rate} The overall value of $\vwO$ has
  contributions from Type Ia SNes and main sequence winds.  The
  relative contributions will depend on both the age of the stellar
  population, $\tage$, and $\Mbh$.  This is illustrated in
  Figure~\ref{fig:vweff}. For black hole masses below $10^7 \Msun$ SNe
  Ia will not contribute to $\vwO$ as the flow time scale is short
  compared to $t_{\rm Ia}$, and the heating is dominated by main
  sequence winds (with $\vwO\simeq 100$ km/s).  On the other hand, for
  $\Mbh \gsim 10^8 \Msun$, heating from Ias becomes important, as
  $\rs>\rIa$.  In this regime, we add the $\vwO$ values from SNe Ias
  and from main sequence winds in quadrature. $\vwO$ from SNe Ia's
  depends on the rate of mass loss ($\eta$), which is in turn a
  function of $t_{\rm age}$. This dependence is why the effective
  $\vwO$ increases with $\tage$ for large $\Mbh$.

@@

For small $\eta$ the effects of cooling will be less pronounced. For
each of our galaxies we give the maximum $\eta$ such that $H/C >$1 in
Table .  Roughly speaking, this is the maximum $\eta$ for which it
would be safe to ignore cooling. $\eta$ depends on the age of the
stellar population, and will realistically vary between $\sim 0.1$ and
$\sim 1$ (see Appendix~\ref{app:eta}).

For $\vwO$=1000 km/s, it is apparent from Figure~\ref{fig:cooling},
that radiative cooling is negligible. On the other hand, for
$\vwO$=200 km/s radiative cooling would be quite important for many of
our profiles and should be included.

@@

\appendix
\clearpage
\section{Analytic Expression for the stagnation radius}
\label{app:rs}
\input{rs}
